### 2020-07-20 paper/paper.tex
second pass through section 3

\documentclass[
  amsfonts,
  amsmath,
  tbtags,
  amssymb,
  aps,
  nobibnotes,
  prl,
  twocolumn,
]{revtex4-2}

\input{preamble}

\begin{document}

%% TITLE
\title{Robust Quantum Optimal Control}

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}

\date{\today}

%% ABSTRACT
\begin{abstract}
  This is a paper about robust quantum optimal control.
\end{abstract}

\maketitle

%% SECTION 1
\section{Introduction}
(Existing work) The leading model of universal
quantum computation is gate-based. The Quantum Optimal Control (QOC)
literature has saught to answer the question "What is the best way to construct a quantum gate?".
Analytic methods
have been demonstrated to construct quantum gates
\cite{zhang2020universal, huang2020engineering, han2020experimental,
  xu2020nonadiabatic, carlini2005quantum}.
Most methods rely on solving the time-dependent
Schroedinger equation analytically or deriving equations of motion
from a suitable Lagrangian.

Some analytical methods
have been presented to address realistic experimental constraints.
Considering the dynamical and geometric phases of state
evolution has led to methods for achieving
robustness to pure dephasing in quantum systems
\cite{xu2020nonadiabitc, han2020experimental, merrill2014progress}.
Additionally, suitable choices of bases have been demonstrated that
allow the experimentalist to infer tradeoffs in longitudinal
and pure dephasing coherence times \cite{huang2020engineering}.
To date, however, no analytical method has been presented
to construct a gate while satisfying all relevant experimental
constraints. Furthermore, most analytic methods have only
been demonstrated in the two-level approximation and become
cumbersome to solve for larger Hilbert space sizes.

Numerical quantum control techniques have developed in parallel
with the analytic frameworks. Numerical quantum control
techniques integrate the time-dependent Schroedinger
equation and use sensitivity analysis to optimize target metrics
of the quantum gate, such as fidelity
\cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
These techniques typically employ zeroth- or first-order
optimizers and make no garauntees of optimality.
Further, they formulate the quantum control
problem as unconstrained or use
projective gradient methods to enforce constraints
\cite{machnes2015gradient}.
The latter approach relies on constructing
bases for constraint spaces--which is not always feasible--and
may hinder convergence.

(This work) We employ the trajectory optimization
literature to formulate the quantum
optimal control problem as a constrained optimization problem.
The trajectory optimization framework allows us to handle arbitrary
constraints on the evolution of the quantum state. Additionally,
we introduce methods for constructing gates robust to longitudinal
relaxation and pure dephasing, the dominant barrier to experimentally
realizing practical quantum computing.

We study the quantum optimal control problem on the fluxonium qubit.
We outline experimentally realistic constraints and map them to
the trajectory optimization framework. For the device we study
we achieve a 1.5x increase in $T_{1}$ times.
We present two methods for achieving robustness to pure dephasing,
and compare them to existing dynamic decoupling methods. We find that 
our methods mitigate dephasing by order X over dynamic decoupling.

(Outline)  First we formulate the quantum optimal control problem
in the trajectory optimization framework. Then, we introduce the dynamics
of the fluxonium device and outline experimental considerations
relevant to gate construction. Next we outline a method for
making the optimization $T_{1}$ aware. Finally, we present
some methods for engineering robustness to pure dephasing and
compare them to existing techniques.

%% SECTION 2
\section{QOC + AL-iLQR}
(QOC Problem Statement) Here we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schroedinger equation (TDSE)
\begin{align}
  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(\textbf{u}(t), t) \ket{\psi(t)}
\end{align}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
depends arbitrarily on the controls ($\textbf{u}(t)$). The controls
are so called because they are the means the experimentalist has to
act on the system.

Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
time steps. Typical integration techniques for the TDSE include
approximating unitary propagators as well as explicit integration methods,
such as Runge-Kutta, of the form
\label{step-approx}
\begin{align}
  \ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}
\end{align}

Quantum optimal control seeks the control
parameters that minimize a functional $J(\textbf{u}(t))$.
In the simplest case the functional is
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(\textbf{u}(t))} \rvert}^{2}$
the infidelity between the inital state evolved
to the final time step ($\ket{\psi_{N}(\textbf{u}(t))}$)
and the target state ($\ket{\psi_{f}}$). In general
$J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional ($\nabla_{u} J(u)$).
They employ a first-order optimizer to minimize $J$ with respect to $u$.

(AL-iLQR Problem Statement) The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the quantum optimal control problem. The quantum optimal control
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. In the LQR formulation
the same functional is evaluated at each time step
\begin{equation}
  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
\end{equation}
where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
at time step $k$ and final state, $u_{k}$ are the controls,
and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
The state is propagated using a dynamics function similar to \ref{step-approx}
$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics.

The advantage of the LQR formulation
is that there exists a dynamic programming algorithm to compute the
optimal update to the controls ($u_{k}$) which minimizes the functional
($J_{\textrm{iLQR}, k}$) for each time step. This algorithm proceeds by deriving a
recurrence relation between time steps $k$ and $k + 1$ for the optimal
feedback law--known as the Ricatti recursion (see Appendix). The
iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
and applies the Ricatti recursion to all time steps on multiple
executions.

In order to incorporate constraints we employ
the augmented Lagrangian method. Constraints are contributions
to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
zero-valued when the constraint is satisfied. The AL-iLQR
method updates the penalty multiplier between
iLQR executions. The penalty multiplier associated with the
constraint functional estimates the Lagrange multiplier of
the constraint. The functional now takes the form
\begin{equation}
  \begin{aligned}
    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
    &+ J_{\textrm{iLQR}}
  \end{aligned}
\end{equation}
$\lambda_{k}$ is a Lagrange multiplier. $I_{\mu_{k}}$ is a penalty matrix
with $\mu_{k}$ along the diagonal. $i$ indicates the $i$-th constraint functional.
$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
\begin{align}
  \lambda_{k_{i}}^{'} &= \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*})\\
  \mu_{k_{i}}^{'} &= \phi \mu_{k_{i}}
\end{align}
where $x^{*}, u^{*}$ are the optimal state and controls from the iLQR execution
and $\phi$ is a hyperparameter. With this updated form of the cost
functional there still exists a recurrence relation to calculate the optimal control
updates, see \cite{howell2019altro}.

%% SECTION 3
\section{QOC on the Fluxonium}
(Fluxonium Device) In the following we study
the quantum optimal control problem on the fluxonium qubit.
In the two-level
approximation the system Hamiltonian takes the form
\label{hamiltonian}
\begin{align}
  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + u(t) \frac{\sigma_{x}}{2}
\end{align}
$h$ is Planck's constant. The control parameter $u$ is experimentally
realized by modulating the flux 
threading the superconducting junction $\Phi_{\textrm{ext}}$.
The conversion is defined by
$u = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} E_{L} \delta \Phi_{\textrm{ext}} / \Phi_{0}$
where $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
offset from the flux frustration point.

In the following we compare the gates found with our methods
to the $X/2, Y/2, Z/2$ gates reported in \cite{zhang2020universal}
for the same device. Note that $Y/2$ and arbitrary $Z$ rotations
are sufficient for universal computation.
%% TODO: list device parameters

(Constraints) We now outline the constraints that we require each gate
to obey.
We require $u(t = 0) = u(t = t_{N}) = 0$.
This constraint ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions.
Furthermore, we require $\int_{0}^{t_{N}} u(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines.
%% TODO: See ref. 28 in Helin paper
We require $-300 \textrm{MHz} \le u(t) \le 300 \textrm{MHz}$
to ensure the two-level approximation \ref{hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\braket{\psi_{f} \lvert \psi_{N}} = 1$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the control
to mitigate AWG ringing.

The equality and inequality constraints on $u$ are handled
with a bound constraint. $\int_{0}^{t_{k}} u(t) dt$ is added
to the augmented state vector $x_{k}$ by incorporating the
integration of $u_{k}$ into the dynamics function $f$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the augmented final state is
reached $x_{N} - x_{f} = 0$. The optimization is performed
over the second derivative of the controls $\frac{d^{2}}{dt^{2}} u(t)$
so that the second derivative, first derivative, zeroth derivative, and
integral of the controls may be penalized.
%% TODO: probably want to list equations for the constraints

($T_{1}$ and $T_{\phi}$ noise)
In addition to optimizing the cost functional to achieve a gate
that obeys experimental constraints and has a high simulated fidelity,
we also want to make the gate robust to noise that affects the experimental
fidelity of the gate. Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
(see Appendix).
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.

Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(\Phi_{\textrm{ext}} = 0.5 \Phi_{0}) \sim 300 \mu\textrm{s}$
where the qubit frequency is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point. Conversely, $T_{1}$ is at a minimum
around the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(0.43 \Phi_{0}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time ($T_{1}$)
over the gate duration
and employ robust control techniques to mitigate
pure dephasing.
%% TODO: See fig. 3 in Helin paper

%% LOH
%% SECTION 4
\section{Robustness to $T_{1}$-type Noise}
(Strategy) We seek to minimize the probability
that the qubit relaxes longitudinally.
To this end we augment the state vector $x_{k}$ with the
longitudinal relaxation probability
\begin{align}
  P_{1}(t_{k}) = \int_{0}^{t_{k}}
  \gamma_{1}(\Phi_{\textrm{ext}}(t)) dt
\end{align}
where $\gamma_{1} = T_{1}^{-1}$. Setting the target
longitudinal relaxation probability to 0 results in
a quadratic cost at each knot point
of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.

The value $T_{1}(\Phi_{\textrm{ext}}(t_{k}))$
is obtained at each knot point by evaluating
a spline interpolant fit to
experimentally obtained data of the form
$\{(\Phi_{\textrm{ext}}, T_{1})\}$.
Calculating $T_{1}$ directly from theoretical
considerations requires many high-dimensional
eigendecompositions, which
is computationally expensive. Additionally,
$T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data increases
the fridge truth of the simulation.

Furthermore, the probability of longitudinal 
relaxation is dependent on the gate duration $t_{N}$.
We allow the optimizer to tune the gate duration by
augmenting the control vector $u_{k}$ with
the time step between knot points $\Delta t_{k}$.
Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. To ensure numerical
integration accuracy is maintained we add a bound
constraint at each knot point
$5\textrm{e-}3 \ \textrm{ns} \le
\Delta t_{k} \le 2\textrm{e-}2 \ \textrm{ns}$.
Note that this bound constraint is allowed to be
broken for intermediate iterations of the optimization,
so we use the absolute value of the time step
$\lvert \Delta t_{k} \rvert$ to ensure that it is non-negative.

(Results) We achieve a factor of 1.5 decrease in the probability
of longitudinal relaxation from the analytic gates we benchmark
against. We simulate the performance of the gates using the
lindblad master equation (see Appendix X). We repeatedly apply the basis gates
and measure the fidelity of the resulting state as a function
of time as shown in Fig. 1.

%% T4.1
%% TODO: Update this table with spline T1 optimizations and calculations.
%% TODO: Maybe best for appendix.
%% TOOD: Maybe add some commentary about why we see more speedup
%% for Z/2 and X/2 than Y/2. In particular I think this is because of the
%% long idle times at the flux frustration point for Z/2 and X/2. It
%% might be a more fair comparison to show gates that have better T1s
%% and as good flux noise mitigation.
\begin{table}
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
    \hline
    Z/2 & 4.615 & 2.576 & 1.792\\
    Y/2 & 2.826 & 1.888 & 1.496\\
    X/2 & 9.562 & 5.080 & 1.882\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

%% F4.1
%% TODO: Update images. Use new gates constructed with
%% spline T1 and use smaller sim times.
\begin{figure}[ht]
  \begin{subfigure}{.33\textwidth}
    %% \includegraphics[width=\linewidth]{assets/00045_spin16.png}
    \label{fig:sub-first}
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    %% \includegraphics[width=\linewidth]{assets/00047_spin16.png}
    \label{fig:sub-second}
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    %% \includegraphics[width=\linewidth]{assets/00049_spin16.png}
    \label{fig:sub-third}
  \end{subfigure}
  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates.}
  \label{fig:fig}
\end{figure}

%% SECTION 5
\section{Robustness to $T_{\phi}$-type Noise}
(Strategy) System parameter deviations arise due to
measurement error and noise injection. For example,
we might measure $\omega_{q}$ inaccurately,
which leads to a deviation from
the the drive frequency $\omega_{d}$.
Further, we may have some temperature dependent gain
in the lines leading to the device, or 1/$f$ noise
and $A(\Phi_{\textrm{ext}})$ comes out with some
$\delta A$ on the flux axis.
Both of these effects lead to dephasing,
uncertain dynamics, and thus erroneous computation.

These problems are typically attacked with dynamic decoupling
sequence where an erroneous rotation is compensated for with
one that cancels the $\delta$. However, we can do better
than the first order variants of these pulses.

We propose two methods for engineering robustness
to system parameter deviations. The first we call the
derivative method. We use the intuition that
making the final state insensitive to changes
in the system parameter is encoded in the derivative
$\partial_{\omega}^{l} \ket{\psi_{N}}$.
In particular, we find that going to the second
derivative is good. The first derivative is already
small, especially for the $\delta A$ time, because,
near the end of optimization, the
optimizer is unable to improve the final state by
changing the flux bias.
The dynamics for propgating the derivative of the state
are found by differentiating the schroedinger equation
dynamics with respect to the parameter of interest.
The scheme boils down to propagating sequential
derivatives of the parameter you want to be
robust to $\partial_{\omega}^{l}$ and penalizing thier norm
${\lvert \partial_{\omega}^{l} \lvert}^{2}$.

The other scheme we call the sampling method. This method
is well-known in the trajectory optimiztion. In this scheme
we propgate an array of states whose dynamics differ
in that the parameter of interest is altered slightly
for each state. In the 3-point sampling method we
propagate $\psi_{\pm}$, $\psi$ where
$\omega_{\pm} = \omega \pm \delta \omega$. Here $\delta$ is
typically taken at standard deviations of the parameter
of interest.

(Results) We compare the simplest schemes
for the derivative and sampling methods to
the CORPSE dynamic decoupling sequence.
They perform X better by Y metric.

We also find that longer gate durations allow for
greater robustness.

%% F5.1
%% TODO: Update this and make sure it is for X/2.
%% I'm not sure if we should include all gates Z/2, Y/2, X/2
%% because I don't know what the dynamic decoupling analogy
%% for Z/2 and Y/2 are. On the other hand we would
%% want to include all of them to show that the benefit
%% is not uniform across gates.
\begin{figure}
  %% \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
  \caption{fidelity v.s. $\omega$ for the $X/2$ gate via
    analytic, 3-sample, 2-derivative (maybe add 5-sample and 3-derivative)}
\end{figure}

%% F5.2
%% TODO: Update this and make sure it is for X/2.
%% Add inset to show high-performing variants at large time steps.
\begin{figure}
  %% \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
  \caption{fidelity at $\omega \pm \sigma_{\omega}$ v.s. time for the
    $X/2$ gate via analytic, 3-sample, 2-derivative
    (maybe add 5-sample and 3-derivative)}
\end{figure}

%% F5.3
%% TODO: Make this figure. My guess is we want a 5-sample, with 2 at \omega_{\pm}
%% and 2 at A_{\pm} as well as 2-derivative of both parameters separately.
\begin{figure}
  %% \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
  \caption{Master equation simulation with $T_{2}$ dissipation
    comparing the gates robust to $\delta \omega_{q}$ and $\delta A$ against
    the analytic gate.}
\end{figure}

%% TODO: Fig: Full ME simulation for best T1 and T\phi pulses
%% v.s. analytic

%% SECTION 6
\section{Discussion}
We have proposed some schemes and they work well.

\appendix
%% APPENDIX 1
\section{Ricatti Recursion}
This will give the reader unfamiliar with trajectory
optimization intuition for how the trajectory optimization
update scheme works and why it is better than
a more naive method.

%% APPENDIX 2
\section{Experiment}
We measure $T_{1}$ using the standard experiment
and $T_{2}$ using te Ramsey experiment. We fit with splines
and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
We measure $\omega_{q}$ and $\sigma_{\omega_{q}}$ using X method.

%% APPENDIX 3
\section{ME Simulation}
We model dissipation using the lindblad master
equation and standard collapse operators for
longitudinal relaxation and dephasing
of a two-level system.

%% APPENDIX 4
\section{Derivative Method}
\begin{equation}
  \begin{aligned}
    \partial_{\omega}^{2} \partial_{t} \psi
    = \frac{-i}{h} [&2 \cdot \frac{\sigma_{z}}{2} \partial_{\omega} \psi\\
      &+ (\omega \frac{\sigma_{z}}{2} + A \frac{\sigma_{x}}{2}) \partial_{\omega}^{2} \psi]
  \end{aligned}
\end{equation}

\bibliography{refs}

\end{document}