### 2020-11-03 paper/main.tex

\documentclass[
  amsfonts,
  amsmath,
  tbtags,
  amssymb,
  aps,
  nobibnotes,
  %% prl,
  twocolumn,
  superscriptaddress,
]{revtex4-2}

\input{preamble}

\begin{document}

%% TITLE
\title{Robust Control of a Fluxonium Qubit}

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zac Manchester}
\affiliation{
  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

\date{\today}

\input{0abstract.tex}

\maketitle

\input{1introduction.tex}
\input{2background.tex}
\input{3fluxonium.tex}
\input{4longitude.tex}
\input{5static.tex}
\input{6stochastic.tex}
\input{7conclusion.tex}
\input{8ack.tex}

\appendix
\input{a1experiment.tex}
\input{a2dissipation.tex}
\input{a3derivative.tex}
\input{a4ctensor.tex}

\bibliography{refs}

\end{document}