### 2021-03-22 paper/2background.tex
jens all edits plus some thomas revisions

\section{Background \label{sec:background}}
In this section, we
review the QOC problem statement
and describe the ALTRO solver \cite{howell2019altro}.
QOC concerns a vector $\mathbf{a}(t)$ of time-dependent control fields
that steer the evolution of a quantum state $\ket{\psi(t)}$.
The evolution of the state is governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(\mathbf{a}(t), t) \ket{\psi(t)}.
  \label{eq:tdse}
\end{equation}
The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system
and the external control fields.
The QOC problem is to find the
controls that minimize a functional $J[\mathbf{a}(t)]$,
which we call the objective.
To make the problem numerically tractable,
the quantum state and controls are discretized into $N$ time steps,
$\ket{\psi(t_{k})} \to \ket{\psi_{k}}$,
and $\mathbf{a}(t_{k})\to \mathbf{a}_{k}$, where $t_{k + 1} = t_{k} + \Delta t$,
and $k \in \{1, ..., N\}$.
In the case of a single state-transfer problem, the objective is
the infidelity of the time-evolved final state $\ket{\psi_{N}}$ and
the intended target state $\ket{\psi_{T}}$,
$J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
Standard QOC solvers compute derivatives of the objective $\nabla J(\mathbf{a})$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is expressed in terms of the cost function
at each time step $\ell_k$, where
$\mathbf{x}_{k}$ is the state vector,
and $\mathbf{u}_{k}$ is the control vector.
The control vector contains all variables that the experimentalist
may manipulate, and the state vector contains all variables that
depend upon those in the control vector.
For example, the control field 
$\mathbf{a}_{k}$ is a component of the control vector $\mathbf{u}_{k}$,
and the quantum state $\ket{\psi_{k}}$ is a component of
the state vector $\mathbf{x}_{k}$.

The dependence
of the state on the controls is encoded in the discrete relationship
$\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
For QOC, the discrete dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$
propagates the quantum state by integrating the TDSE with Runge-Kutta methods \cite{jorgensen2011numerical}
or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
shillito2020fast}. \todo{Jens wanted TDSE related before ddf}

Additional constraints imposed on the augmented controls and augmented states
are formulated as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$ or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$. \textcolor{blue}{[How to understand the $k$? labels here?]}
The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints, and equalities and inequalities are understood component-wise. Each constraint function's \emph{violation} is defined as the magnitude of the maximum deviation: 
$\textrm{max}(\mathbf{g}(\cdot), 0)$ or $\textrm{max}(\abs{\mathbf{h}(\cdot)})$
where $\textrm{max}(\cdot)$ and $\abs{\cdot}$ are understood to act element-wise. \textcolor{blue}{[Is this the right place for this definition in the text. Seems unmotivated/out of place right now?]}
Stated concisely, the trajectory optimization problem is
\begin{mini!}[2]
  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\ell_{N}(\mathbf{x}_N) + \sum_{k=1}^{N-1}
    \ell_{k}(\mathbf{x}_k,\mathbf{u}_k)}{}{} \label{eq:gcostfun}
    \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \quad \forall \ k} \label{eq:gdyncon}
    \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \quad \forall \ k} \label{eq:ineqcon}
    \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \quad \forall \ k.} \label{eq:eqcon}
\end{mini!}
\textcolor{blue}{[Can't we just define $\mathbf{u}_N=0$ to simplify the expression in (2a)?]}

Standard techniques for solving \eqref{eq:gcostfun}-\eqref{eq:eqcon} typically
fall into two categories: direct methods \cite{Hargraves87, kelly2017introduction}
and indirect methods \cite{betts1998survey}. For indirect methods,
the augmented controls are the \emph{decision variables}, i.e., the
variables the optimizer adjusts to solve the problem.
The augmented states are obtained from the augmented controls using the discrete dynamics function,
and are then used to evaluate derivatives of the cost functions \eqref{eq:gcostfun}.
Then, the derivative information is employed to update the augmented controls.
This approach is taken by standard QOC solvers such as GOAT \cite{machnes2015tunable},
GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
Conversely, direct methods treat both the augmented controls and the augmented states as decision
variables. In addition to minimizing the cost functions, the optimizer uses derivative information
for the discrete dynamics function
to satisfy the dynamics constraint \eqref{eq:gdyncon} to a specified tolerance.
In this sense, the TDSE \eqref{eq:tdse} is a constraint that may be violated
for intermediate steps of the optimization, where the states need not be physical.
The direct approach lends itself to a nonlinear program formulation, for which
a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.

Recent state-of-the-art solvers, such as ALTRO,
 combine the indirect and direct method in a two-stage approach.  Specifically, ALTRO first employs the indirect method as implemented by an
iterative linear-quadratic regulator (iLQR) algorithm
\cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
\cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}. In the second and final stage to the solution, the direct-method approach is utilized in the form of a projected Newton method \cite{bertsekas1982projected, rao1998application}. We next provide a more detailed summary of these two stages.

iLQR is an indirect method
for solving the dynamically constrained
trajectory optimization problem \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
and its update procedure is based on the differential-dynamic-programming approach
\cite{mayne1966a}. \textcolor{blue}{[Start with the familiar or the understandable explanation. The imposing sounding names can be given afterwards. (Purely psychology.)]}
First, iLQR uses an initial guess for the augmented controls to obtain the augmented
states with the discrete dynamics function.
iLQR then constructs quadratic models for each cost function using
their zeroth-, first- and second-order derivatives in a Taylor expansion
about the current augmented controls and augmented states.
These models are used to derive a recurrence relation between time steps \textcolor{blue}{[for some quantity (which?) at consecutive time steps]}
which gives the locally optimal update for the augmented controls.
Finally, a line search \cite{zhang2006global}
is performed in the direction of this update to ensure a
decrease in the objective \eqref{eq:gcostfun}. \textcolor{blue}{[again, I suspect that (2a) is not the ``objective"; rather the objective is what I know as cost function, which would exclude the minimization that is part of (2a).]} This procedure
is repeated until convergence is reached.

While indirect solvers like iLQR are computationally efficient and maintain high accuracy for the discrete dynamics throughout the optimization, they typically cannot handle
nonlinear equality and inequality
constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
Projected gradient methods are a preferred approach
to handle such constraints
\cite{clarkson2010coresets, hauswirth2016projected, morzhin2019minimal,
  nikolskii2007convergence}. Unfortunately, within the indirect framework,
they can only be used for constraints on the augmented controls,
not the augmented states. \textcolor{blue}{[Seems like my edit attempts and understanding of the two-stage process are still not appropriate. On the previous page, it seemed like the projected Newton method was the final stage and a direct method at that. However, here it sounds now like it is not good enough and we need yet another method? I'm confused.]}
Another technique, which is popular for QOC \cite{heeres2017implementing, leung2017speedup,
reinhold2019controlling},
is to add the constraint functions to the objective
\eqref{eq:gcostfun}. \textcolor{blue}{[(2a) is of the form $\text{min}_x L(x)$ do you really mean to add to the minimum? Or only to the cost function $L(x)$?]}
This strategy does not guarantee that the constraints
are satisfied as the solver trades
minimization of the cost functions and constraint functions against each other.
ALM remedies this issue by adaptively adjusting a Lagrange multiplier estimate
for each constraint function to ensure the constraints are satisfied.
ALM adds terms that are linear and quadratic in the constraint functions
to the objective. Then, the new objective is minimized with
iLQR. If the solution obtained with iLQR does not satisfy the constraints,
the prefactors for the constraint terms in the objective are increased
intelligently and the procedure is repeated.

ALM converges superlinearly, but poor numerical conditioning may lead
to small decreases in the constraint violations near the locally optimal solution
\cite{bertsekas1996constrained}.
To address this shortcoming, ALTRO
projects the solution from the ALM stage onto the constraint manifold using
a (direct) projected Newton method, achieving ultra-low
constraint violations $\sim 10^{-8}$.
For more information on the details of the ALTRO
solver, see \cite{howell2019altro, Jackson2020altroc}.

As opposed to standard QOC solvers, ALTRO
can satisfy constraints
on both the augmented controls and the augmented states to tight tolerances.
This advantage is crucial for this work, where multiple medium-priority cost functions
are minimized subject to many high-priority constraints.