### 2021-03-22 paper/3fluxonium.tex

\section{QOC for the Fluxonium \label{sec:fluxonium}}
In the following, we optimize quantum gates
(unitary transformations) \textcolor{blue}{[for a physics journal, you will want as much handholding and explanation for the robotics community language as possible; bu contrast, clarifying that a quantum gate is a unitary tranformation seems out of place here.]} for the superconducting fluxonium qubit -- a promising
building block for quantum computers due to its high
coherence times
\cite{earnest2018realization, lin2018demonstration,
  manucharyan2009fluxonium, nguyen2019high,
  zhang2020universal}.
In this section, we use the trajectory optimization
formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
to define the optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
which we extend in subsequent sections to account
for experimental error channels.
To high accuracy, we can approximate the fluxonium Hamiltonian near the flux-frustration
point by a two-level system:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}.
  \label{eq:hamiltonian}
\end{align}
Here, $f_{q}$ is the qubit frequency at the flux-frustration point,
$a(t)$ is the control governing the flux offset from the flux-frustration point,
$h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
are Pauli matrices. Although the coherent dynamics can be described based on this two-level
system model, our noise model, experimental constraints, and system parameters
consider the full system, and they are representative of the fluxonium
presented in \cite{zhang2020universal}.

First, we outline the constraints for the fluxonium gate problem.
Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
 $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
\eqref{eq:istate_con}
where the superscript $i \in \{0, 1\}$ is the state index,
and the subscript labels the time step -- here, $k = 1$.
The states at the final time step are constrained to be
the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv
U \ket*{\psi^{i}_{1}} \ \forall \ i$
\eqref{eq:tstate_con} with $U = X/2, Y/2, Z/2$ denoting the target gate.
Furthermore, we impose the normalization constraint
${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
\eqref{eq:statenorm_con}
to ensure the solver does not take advantage of discretization errors in numerical integration.
To refer to the discrete moments of the flux, we introduce the notation
$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
$a(t_{k}) \equiv a_{k}$,
$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$. \textcolor{blue}{[Why is this a partial rather than a single-variable derivative?]}
We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
\eqref{eq:znf_con}
which mitigates the inductive drift ubiquitous in flux-bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
\eqref{eq:amp_con}.
Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con} \textcolor{blue}{[Thomas, this would actually be consistent with my suggestion to set $\mathbf{u}_N$ to zero?]}
so the gates may be concatenated arbitrarily. Additionally,
we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
a maximum violation of $\sim 10^{-8}$.

The augmented control and augmented state are:
\begin{equation}
  \begingroup
  \renewcommand*{\arraystretch}{1.3}
  \mathbf{u}_{k} = \partial^{2}_{t} {a}_{k} , \quad
  \mathbf{x}_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
  \endgroup
  \label{eq:astatecontrols}
\end{equation}
(We maintain vector notation for $\mathbf{u}_k$ in anticipation of augmentation with additional components in Section \ref{sec:longitude}.)
The time evolution of these variables is governed by 
 coupled first-order differential equations
which are encoded in the discrete dynamics function \eqref{eq:dyn_con}.
We integrate the states according to the TDSE \eqref{eq:tdse} based on the
fluxonium Hamiltonian \eqref{eq:hamiltonian}.
Since the ALTRO implementation we use does not currently
support complex numbers, we follow \cite{leung2017speedup} in  representing the states
via the isomorphism $\mathcal{H}(\mathbb{C}^{n})
\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
\begin{equation}
  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}.
  \label{eq:isomorphism}
\end{equation}

The cost function at each time step is
$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k} - \mathbf{x}_{f})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{f}) + \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply. \textcolor{blue}{[``We supply" -- sure, but not very helpful. How are they defined/constructed specifically?]}
The $Q_{k}$ term
penalizes deviations from the target augmented state $\mathbf{x}_{f}$,
which is given by the constraints we have imposed on
$\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
$\partial_{t} a_{f} = 0$. \textcolor{blue}{[Should this have been $a_N$ rather than $a_f$?]}
The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
smoothing the flux to mitigate high-frequency AWG transitions.
Stated succinctly, the optimization problem takes the form:
\begin{mini!}[2] 
  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_f)}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{f})
    + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, u_k) \quad \forall \ k}  \label{eq:dyn_con}
  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
    \quad \forall \  i} \label{eq:tstate_con}
  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
      \rvert}^{2} = 1 \quad \forall \ i, k} \label{eq:statenorm_con}
  \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \quad \forall \ k} \label{eq:amp_con}
  \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0}. \label{eq:ic_con}
\end{mini!}

Next, we remark on our problem formulation.
We penalize the squared error of the final state and the target
state \eqref{eq:costfun} \textcolor{blue}{[referencing (6a) is puzzling here unless you specify $R_k$; how is ``squared error" different from infidelity?]}, rather than their infidelity,
because the Hessian of the squared-error cost function is diagonal -- which
makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
which requires a metric that is sensitive to global phases for the initial
states $\ket{0}$ and $\ket{1}$.
We penalize the squared error at all time steps
because it benefits the iLQR solving stage \cite{Jackson2020altroc},
not to incentivize early achievement \textcolor{blue}{[NOT to incentivize or to incentivize??]}
of the desired gate \cite{leung2017speedup}.

In addition to the target-state cost function, we
impose a target-state constraint \eqref{eq:tstate_con} which  requires the final state to match the target state, including its global phase,
up to our chosen maximum constraint violation $\sim 10^{-8}$.
If we did not impose this constraint, the optimizer would be
allowed to sacrifice the closed-system gate error to achieve better
performance on the other cost functions, which is undesirable.
To enforce a constraint in standard QOC frameworks,
the prefactor for the constraint function is manually increased
between separate optimization instances until the constraint is satisfied
\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling},
which becomes infeasible for more than one constraint.
ALM automates these prefactor updates to find
a solution that satisfies all of the given constraints.
Hence, ALTRO's ability to handle multiple constraints makes it
an attractive solver for QOC problems.

In extraordinarily difficult cases of
QOC \cite{abdelhafez2020universal}, it may be impossible
to obey the physics of the system and achieve the desired gate;
equivalently, the dynamics constraint \eqref{eq:dyn_con}
and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
In this case, the prefactors for the constraint function terms
in the ALM objective will tend to infinity -- leading to numerical instability -- and the
optimization will not converge. To maintain a constrained approach in this situation,
the maximum constraint violation for the target-state constraint can be raised
to a level commensurate with the minimum acceptable gate error.

Finally, for the indirect ALM-iLQR solving stage,
the state at each time step is obtained by integrating
the TDSE with the discrete dynamics function,
so the dynamics constraint \eqref{eq:dyn_con} is satisfied by construction.
This is not the case for the direct projected Newton stage, where the states
are free parameters that are adjusted to satisfy the TDSE.
Although the final solution's deviation from the TDSE is never more than the maximum constraint violation,
we explicitly integrate the TDSE when reporting gate errors to ensure accuracy.
Exploring the benefit of direct optimization approaches for QOC is an interesting direction for future work.