### 2021-03-22 paper/8ack.tex

\begin{acknowledgments}
  We thank Helin Zhang for experimental assistance
  and Taylor Howell, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
  This work is funded in part by EPiQC, an NSF Expedition in Computing, under grant CCF-1730449.
  This work was supported by the Army Research Office under Grant No. W911NF1910016.
  This work was made possible by many open source software projects,
  including but not limited to:
  Altro.jl \cite{howell2019altro},
  DifferentialEquations.jl \cite{rackauckas2017differentialequations},
  Distributions.jl \cite{besancon2019distributions},
  ForwardDiff.jl \cite{revelsLubinPapamarkou2016},
  Matplotlib \cite{hunter2007matplotlib},
  NumPy \cite{harris2020array},
  and Zygote.jl \cite{innes2018don}.
\end{acknowledgments}