### 2021-03-26 paper/a4time.tex

\section{Computational Performance \label{appendix:time}}
In this section we provide runtimes for our optimizations. The runtimes for
the base optimization in Sec. \ref{sec:fluxonium},
the depolarization optimization in Sec. \ref{sec:longitude},
and the robust optimizations in Sec. \ref{sec:static}
are presented in Table \ref{tab:time}
for a $Z/2$ gate at gate times which are multiples of $1/4f_{q} \sim 18$ns.
We performed optimizations on a single core of
an AMD Ryzen Threadripper 3970X 32-Core Processor.
Future work will parallelize the robustness methods using GPUs
\cite{leung2017speedup},
which will enable fast optimizations on high-dimensional Hilbert spaces.

\begin{table}[H]
  \centering
  \begin{tabular} {| c | c | c | c | c |}
    \hline
    \backslashbox{Method}{$t_{N} (\textrm{ns})$} & $18$ & $36$ & $72$\\
    \hline
    Base & $0.155 \pm 0.008$ & $7.0 \pm 0.4$ & $15.9 \pm 0.8$\\
    \hline
    Depol. & $1.69 \pm 0.08$ & - & -\\
    \hline
    S & $1.77 \pm 0.09$ & $48 \pm 2$ & $280 \pm 10$\\
    \hline
    U & $75 \pm 4$ & $340 \pm 20$ & $400 \pm 20$\\
    \hline
    D1 & $6.1 \pm 0.3$ & $27 \pm 1$ & $65 \pm 3$\\
    \hline
    D2 & $15.7 \pm 0.8$ & $17.3 \pm 0.9$ & $54 \pm 3$\\
    \hline
  \end{tabular}
  \caption{
    Average runtimes in seconds for $Z/2$ optimizations
    using the base, depolarization, sampling (S),
    unscented sampling (U), and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
  }
  \label{tab:time}
\end{table}