### 2021-03-26 paper/main.tex

\documentclass[
  amsfonts,
  amsmath,
  amssymb,
  pra,
  twocolumn,
  superscriptaddress,
]{revtex4-2}

\input{preamble}

\begin{document}

\input{0abstract.tex}
\input{1introduction.tex}
\input{2background.tex}
\input{3fluxonium.tex}
\input{4longitude.tex}
\input{5static.tex}
\input{6stochastic.tex}
\input{7conclusion.tex}
\input{8ack.tex}

\appendix
\input{a1longitude.tex}
\input{a2unscented.tex}
\input{a3derivative.tex}
\input{a4time.tex}

\bibliography{refs}

\end{document}